---
title: "non-linear-least-squares"
author: "Carmen Galaz-García"
date: "3/5/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)

library(tidyverse)
library(broom)
```


## Preparing data
```{r}

raw_fish <- read_csv("fish_catch.csv")

fish <- raw_fish %>% 
  slice(5:67) %>% 
  select(1,2) 

colnames(fish)[1] <- "year"
colnames(fish)[2] <- "wild_catch"

fish <- fish %>% 
  mutate(year= as.integer(year)) %>% 
  mutate(year= year-1950) %>% 
  mutate(wild_catch= as.numeric(wild_catch)) 

``` 

### Data exploration

```{r}
# Data exploration
ggplot(data = fish, aes(x = year, y = wild_catch)) +
  geom_point() +
  theme_minimal() +
  labs(x = "years since 1950 ", 
       y = "world wild catch (million tons)")
  
```


```{r}
# Find Estimates exp_fish <- fish %>% 
  filter(year < 20) %>% 
  mutate(ln_catch = log(wild_catch))
  
# Model linear to get *k* estimate (the slope of this linear equation is an estimate of the growth rate constant):
lm_k <- lm(ln_catch ~ year, data = exp_fish)
lm_k
```

The data seems to follow a logistic growth (though for the first 20 years it seems to be quite linear). This is if $t$ is the years since 1950 then the world wild fish catch (in millions of tons is) $P$ as a function of $t$ could be modeled by
$P(t) = \frac{K}{1+Ae^{-kt}}$ where

- $K$ is the carrying capacity, in this case the total of fish tons available for fishing in the wild,
- $A = \frac{K-P_0}{P_0}$, where $P_0$ is the catch at time 0, and
- $k$ is the growth rate constant. 

We estimate these parameters to be 
$K\sim 92$ and
$P_0 \sim 17$, which gives $A \sim 4.41$, and
$k \sim 0.058$, 




## Nonlinear least squares
```{r}


nls_catch <- nls(wild_catch ~ K/(1 + A*exp(-r*year)),
              data = fish,
              start = list(K = 92, A = 4.41, r = 0.058),
              trace = FALSE
              )
summary(nls_catch)
```
```{r}
model_out <- broom::tidy(nls_catch) 
K_est <- model_out$estimate[1] 
A_est <- model_out$estimate[2] 
k_est <- model_out$estimate[3] 
```

Based on a non-linear least-squares model, the optimized parameters are
`r K =K_est` millions of tons of wild fish caught,
`r A =A_est`  and
`r k =k_est` millions of tons of wild fish caught per year. 


### Model visualization

```{r}
# Make predictions for the population at all of those times (time) in the original df: 
catch_predict <- predict(nls_catch)

# Bind predictions to original data frame:
catch_complete <- data.frame(fish, catch_predict)

# Plot them all together:
ggplot(data = catch_complete, aes(x = year, y = wild_catch)) +
  geom_point(col='blue4', alpha=0.5) +
  geom_line(aes(x = year, y = catch_predict)) +
  theme_minimal()+
  labs( title= "",
    x = "years since 1950 ", 
       y = "world wild fish catch (million tons)")
```

$## Data source

Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), ​Global Capture Production​ and Global Aquaculture Production​, electronic databases, at www.fao.org/fishery/topic/16140/en.
