---
title: "bootstrapping"
author: "Carmen Galaz-García"
date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(boot)
```

```{r}
# DATA CREATION
# 1 = reported experiencing exclusionary, offensive, hostile or intimidating conduct
# 0 = no reporting
students <-rep(c(1,0),times=c(22,14))

# FUNCTION ON DATA
proportion <- function(x,i){sum(x[i])/36}

# BOOTSTRAPPED PROPORTIONS
studentsboot <- boot(students, 
                     statistic = proportion,
                     R=100000)

```

```{r}
# HISTOGRAM

df_studentsboot <- data.frame( bs_proprtion = studentsboot$t)

# Histogram of 10k bootstrap sample means:
ggplot(data = df_studentsboot, aes(x = bs_proprtion)) +
  geom_histogram(bins=25) +
  theme_minimal()+
  labs(title= "Histogram of bootstrapped proportions")


```


```{r}
CI <- boot.ci(studentsboot, conf = 0.95)
```

A final statement (in text) describing the bootstrapped CI in the context of the data (suitable for publication).

Blue whale mean weight is 87.2 tons (n = 11), with a bootstrapped 95% confidence interval of [77.1, 98.0] tons (n = 10,000 bootstrap samples).
The proportion of UCSB nonbinary/genderqueer community members who responded that they had personally experienced “exclusionary, offensive, hostile or intimidating conduct” is $12/36 \simeq 0.61$ (n=36), 
with a bootstrapped 95% confidence interval of 
`r `
